API design for machine learning software: experiences from the scikit-learn project

### 2013-06-27 paper.tex
trim introduction
It ran from page 1, bottom, to page 3, top; now stops at page 2, bottom.

The \sklearn project\footnote{\url{http://scikit-learn.org}}
\citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a general-
purpose high-level language. Among other things, it includes classical learning
algorithms, model evaluation and selection tools, as well as
preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

From its core, the library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a numeric array data type and proper
array computing primitives, while SciPy \citep{varoquaux2013scipy} extends it
further by making common numerical operations available, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. The \textit{scikits} were created to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn for machine learning and 
\textit{scikit-image}\footnote{\url{http://scikit-image.org}} for image
processing.

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g., machine learning,
natural language processing, neuroscience, astrophysics, etc.). The project also benefits
from many occasional contributors proposing small code or documentation fixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Complete documentation and examples are required for all features,
and major modifications must pass code review by at least two
developers not involved in the implementation of the proposed change.

Over the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gauged from various indicators such as the hundreds
of citations in scientific publications, successes in various machine learning
challenges (e.g., Kaggle), or significant statistics derived from our
repositories and mailing list.  At the time of writing, the project is watched
by 1365 people and forked 693 times on GitHub; the mailing list receives more
than 300 mails per month; version control logs
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 unique contributors to the codebase and the online documentation
website receives 37,000 unique visitors and 295,000 page views per month.

In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
benchmarked against several competitors. In this paper, we instead focus on an
in-depth discussion and analysis of the design choices made when building the
library. We first present in section \ref{sec:core-api} the central application
programming interface (API) and then describe, in section \ref{sec:advanced-api},
advanced API mechanisms built from the core interface.
Section \ref{sec:implementation} briefly describes the implementation.
Section \ref{sec:comparison} then
compares \sklearn to other major projects in terms of API.
Section \ref{sec:future_work} outlines some of the work to be done before
the \sklearn 1.0 release.
Finally, we summarize and conclude in section \ref{sec:conclusions}.

>>>>

The \sklearn project\footnote{\url{http://scikit-learn.org}}
\citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a
general-purpose high-level language. Among other things, it includes classical learning
algorithms, model evaluation and selection tools, as well as
preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

From its core, the library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. The \textit{scikits} were created to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn for machine learning and 
\textit{scikit-image}\footnote{\url{http://scikit-image.org}} for image
processing.

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

\sklearn's popularity can be gauged from various indicators such as the hundreds
of citations in scientific publications, successes in various machine learning
challenges (e.g., Kaggle), and statistics derived from our
repositories and mailing list.  At the time of writing, the project is watched
by 1365 people and forked 693 times on GitHub; the mailing list receives more
than 300 mails per month; version control logs
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 unique contributors to the codebase and the online documentation
receives 37,000 unique visitors and 295,000 pageviews per month.

In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
benchmarked against several competitors. In this paper, we instead present an
in-depth analysis of design choices made when building the
library. We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section \ref{sec:implementation} briefly describes the implementation.
Section \ref{sec:comparison} then
compares \sklearn to other major projects in terms of API.
Section \ref{sec:future_work} outlines some of the work to be done before
a \sklearn 1.0 release.
Finally, we summarize and conclude in section \ref{sec:conclusions}.
